Lineage Analysis of CRISPR and shRNA Screens

### 2018-08-29 lineages_18Q3.Rmd
Wilcox tests: CRISPR, gene-tissue grouping

Load libraries.
```{r}
library(NMF)
library(ggsignif)
library(ggpubr)
library(rowr)
library(data.table)
library(CePa)
library(plyr)
library(tidyverse)
library(magrittr)
library(ggrepel)
library(matrixStats)
library(parallel)
library(kableExtra)
library(gridExtra)
library(broom)
library(glmnet)
library(devtools)
library(reshape2)
library(caret)
library(bsselectR)
library(ComplexHeatmap)
library(circlize)
```

@@

  # Gene expression
  lo_ge <- range(gene$RPKM[!is.na(gene$RPKM)])[1]
  hi_ge <- range(gene$RPKM[!is.na(gene$RPKM)])[2]
  mid_ge <- (hi_ge - lo_ge) / 2
  plot_ge <- ggplot(data = gene, aes(x = RPKM, y = Score, color = Mutation_Status)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_ge + mid_ge) / 2, (hi_ge - mid_ge) / 2 + mid_ge), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Gene Expression (RPKM)"),
         title = "Gene Expression",
         subtitle = "r: Pearson correlation coeffcient")

@@

# Convert log2 ratios (log2(CN/2)) to CN
cn[2:ncol(cn)] <- lapply(cn[2:ncol(cn)], function(x) 2 * (2 ^ x))

@@

# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)

saveRDS(crispr_data, "./data_munging/rds/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./data_munging/rds/crispr_data_18Q3.rds")
crispr_ccl <- data.frame("Broad_ID" = crispr_data$Broad_ID)
write.table(crispr_data[, c("Hugo_Symbol", "Broad_ID", "Score", "Primary.Disease", "primary_tissue", "Mutation_Status")], file = "~/Desktop/crispr_data_select.tsv", quote = FALSE, sep = "\t")
```

@@

@@

# Gene expression
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))[, c("Hugo_Symbol", "Broad_ID", "RPKM")]
chm_ge_grid <- merge(crispr_grid, ge_filt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
chm_ge <- reshape(chm_ge_grid, idvar = "Hugo_Symbol", timevar = "Broad_ID", direction = "wide")
chm_ge_rows <- chm_ge$Hugo_Symbol
chm_ge$Hugo_Symbol <- NULL
chm_ge_cols <- colnames(chm_ge) %>% gsub("RPKM.", "", .)
chm_ge <- matrix(as.numeric(unlist(chm_ge)), nrow = nrow(chm_ge))
rownames(chm_ge) <- chm_ge_rows
colnames(chm_ge) <- chm_ge_cols
saveRDS(chm_ge, "./data_munging/rds/crispr_heatmap_ge.rds")

@@

# Copy number
cn_filt <- filter(cn_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))
chm_cn_grid <- merge(crispr_grid, cn_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
chm_cn <- reshape(chm_cn_grid, idvar = "Hugo_Symbol", timevar = "Broad_ID", direction = "wide")
chm_cn_rows <- chm_cn$Hugo_Symbol
chm_cn$Hugo_Symbol <- NULL
chm_cn_cols <- colnames(chm_cn) %>% gsub("Copy_Number.", "", .)
chm_cn <- matrix(as.numeric(unlist(chm_cn)), nrow = nrow(chm_cn))
rownames(chm_cn) <- chm_cn_rows
colnames(chm_cn) <- chm_cn_cols
saveRDS(chm_cn, "./data_munging/rds/crispr_heatmap_cn.rds")
```

@@

@@

chm_plot_cgc <- Heatmap(chm_scores_cgc, name = "CERES Score",
        bottom_annotation = HeatmapAnnotation(df = chm_annot),
        bottom_annotation_height = unit(5, "in"),
        row_title = "Genes", column_title = "Cancer Cell Lines (Broad IDs)",
        col = colorRamp2(c(min(chm_scores_cgc), 0, max(chm_scores_cgc)), c("purple4", "white", "seagreen4")),
        na_col = "black",
        row_title_gp = gpar(fontsize = 60, fontface = "bold"),
        row_names_gp = gpar(fontsize = 10),
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_names_gp = gpar(fontsize = 8),
        row_dend_reorder = TRUE,
        column_dend_reorder = TRUE,
        clustering_distance_rows = "euclidean",
        clustering_distance_columns = "euclidean",
        clustering_method_rows = "complete",
        clustering_method_columns = "complete",
        column_dend_height = unit(4, "in"),
        row_dend_width = unit(4, "in"),
        width = 3)

@@

chm_ge_plot_cgc <- Heatmap(chm_ge_cgc, name = "Gene Expression (RPKM)",
        col = colorRamp2(c(min(chm_ge_cgc[!is.na(chm_ge_cgc)]), 0.25, max(chm_ge_cgc[!is.na(chm_ge_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Gene Expression",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_cn_plot_cgc <- Heatmap(chm_cn_cgc, name = "Copy Number",
        col = colorRamp2(c(min(chm_cn_cgc[!is.na(chm_cn_cgc)]), 2, max(chm_cn_cgc[!is.na(chm_cn_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Copy Number",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_mut_plot_cgc <- Heatmap(chm_mut_cgc, name = "Mutation Status",
        na_col = "black",
        col = c("cyan3", "darkorchid"),
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Mutation Status",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)

@@

chm_ge_plot_cgc <- Heatmap(chm_ge_cgc, name = "Gene Expression (RPKM)",
        col = colorRamp2(c(min(chm_ge_cgc[!is.na(chm_ge_cgc)]), 0.25, max(chm_ge_cgc[!is.na(chm_ge_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Gene Expression",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_cn_plot_cgc <- Heatmap(chm_cn_cgc, name = "Copy Number",
        col = colorRamp2(c(min(chm_cn_cgc[!is.na(chm_cn_cgc)]), 2, max(chm_cn_cgc[!is.na(chm_cn_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Copy Number",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_mut_plot_cgc <- Heatmap(chm_mut_cgc, name = "Mutation Status",
        na_col = "black",
        col = c("cyan3", "darkorchid"),
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = FALSE, show_column_names = FALSE,
        column_title = "Mutation Status",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)

@@

@@

@@

```{r}
makeCRISPRlinplot("HMGA1")
test <- filter(crispr_signif_lineage, Hugo_Symbol == "HMGA1")
```

>>>>

Load libraries.
```{r}
library(NMF)
library(ggpubr)
library(rowr)
library(data.table)
library(CePa)
library(plyr)
library(tidyverse)
library(magrittr)
library(matrixStats)
library(parallel)
library(kableExtra)
library(gridExtra)
library(broom)
library(glmnet)
library(devtools)
library(reshape2)
library(caret)
library(bsselectR)
library(ComplexHeatmap)
library(circlize)
```

@@

  # Gene expression
  lo_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[1]
  hi_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[2]
  mid_ge <- (hi_ge - lo_ge) / 2
  plot_ge <- ggplot(data = gene, aes(x = RPKM_log2, y = Score, color = Mutation_Status)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_ge + mid_ge) / 2, (hi_ge - mid_ge) / 2 + mid_ge), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Gene Expression [log2(RPKM)]"),
         title = "Gene Expression",
         subtitle = "r: Pearson correlation coeffcient")

@@

# Convert log2 ratios (log2(CN/2)) to CN
# cn[2:ncol(cn)] <- lapply(cn[2:ncol(cn)], function(x) 2 * (2 ^ x))

@@

# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)

saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../crispr_lineages_giant_files/crispr_data_18Q3.rds")
crispr_ccl <- data.frame("Broad_ID" = crispr_data$Broad_ID)
write.table(crispr_data[, c("Hugo_Symbol", "Broad_ID", "Score", "Primary.Disease", "primary_tissue", "Mutation_Status")], file = "~/Desktop/crispr_data_select.tsv", quote = FALSE, sep = "\t")
```

@@

### Grouped by gene and tissue

```{r, eval=TRUE}
crispr_signif_tissue <- compare_means(Score ~ Mutation_Status, group.by = c("Hugo_Symbol", "primary_tissue"), data = crispr_data, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_tissue <- adj_signif(crispr_signif_tissue)
# crispr_signif_tissue <- mutate(crispr_signif_tissue, primary_tissue = reorder(primary_tissue, p.adj, mean))
crispr_signif_tissue <- crispr_signif_tissue[order(crispr_signif_tissue$p),]
saveRDS(crispr_signif_tissue, "./data_munging/rds/crispr_signif_tissue.rds")
```

```{r}
crispr_signif_tissue <- readRDS("./data_munging/rds/crispr_signif_tissue.rds")

knitr::kable(filter(crispr_signif_tissue, p < 0.1)[, c("Hugo_Symbol", "primary_tissue", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CRISPR Screen: Wilcoxon Test Results Comparing Mutant and Wildtype Cell Lines by Primary Tissue, p < 0.1 (Benjamini-Hochberg-corrected p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% footnote(general = "Wilcoxon test Benjamini-Hochberg-corrected p-values:\n*: p <= 0.05\n**: p <= 0.01\n***: p <= 0.001\n****: p <= 0.0001") %>% scroll_box(width = "900px", height = "450px")
```

@@

# Gene expression
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))[, c("Hugo_Symbol", "Broad_ID", "RPKM")]
ge_filt$RPKM <- log2(ge_filt$RPKM + 0.0001)
chm_ge_grid <- merge(crispr_grid, ge_filt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
chm_ge <- reshape(chm_ge_grid, idvar = "Hugo_Symbol", timevar = "Broad_ID", direction = "wide")
chm_ge_rows <- chm_ge$Hugo_Symbol
chm_ge$Hugo_Symbol <- NULL
chm_ge_cols <- colnames(chm_ge) %>% gsub("RPKM.", "", .)
chm_ge <- apply(chm_ge, 2, as.double)
chm_ge <- apply(chm_ge, 2, function(x) (x - mean(x, na.rm = TRUE)) / var(x, na.rm = TRUE))
rownames(chm_ge) <- chm_ge_rows
colnames(chm_ge) <- chm_ge_cols
saveRDS(chm_ge, "./data_munging/rds/crispr_heatmap_ge.rds")

@@

# Copy number
cn_filt <- filter(cn_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol))
chm_cn_grid <- merge(crispr_grid, cn_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
chm_cn <- reshape(chm_cn_grid, idvar = "Hugo_Symbol", timevar = "Broad_ID", direction = "wide")
chm_cn_rows <- chm_cn$Hugo_Symbol
chm_cn$Hugo_Symbol <- NULL
chm_cn_cols <- colnames(chm_cn) %>% gsub("Copy_Number.", "", .)
chm_cn <- apply(chm_cn, 2, as.double)
chm_cn <- apply(chm_cn, 2, function(x) (x - mean(x, na.rm = TRUE)) / var(x, na.rm = TRUE))
rownames(chm_cn) <- chm_cn_rows
colnames(chm_cn) <- chm_cn_cols
saveRDS(chm_cn, "./data_munging/rds/crispr_heatmap_cn.rds")
```

@@

```{r}
mean(chm_cn_cgc[!is.na(chm_cn_cgc)])
```

@@

chm_plot_cgc <- Heatmap(chm_scores_cgc, name = "CERES Score",
        bottom_annotation = HeatmapAnnotation(df = chm_annot),
        bottom_annotation_height = unit(5, "in"),
        row_title = "Genes", column_title = "Cancer Cell Lines",
        col = colorRamp2(c(min(chm_scores_cgc), 0, max(chm_scores_cgc)), c("purple4", "white", "seagreen4")),
        na_col = "black",
        row_title_gp = gpar(fontsize = 60, fontface = "bold"),
        row_names_gp = gpar(fontsize = 10),
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_names_gp = gpar(fontsize = 8),
        row_dend_reorder = TRUE,
        column_dend_reorder = TRUE,
        clustering_distance_rows = "euclidean",
        clustering_distance_columns = "euclidean",
        clustering_method_rows = "complete",
        clustering_method_columns = "complete",
        column_dend_height = unit(4, "in"),
        row_dend_width = unit(4, "in"),
        width = 3)

@@

chm_ge_plot_cgc <- Heatmap(chm_ge_cgc, name = "Gene Expression [log2(RPKM)]",
        col = colorRamp2(c(min(chm_ge_cgc[!is.na(chm_ge_cgc)]), 0, max(chm_ge_cgc[!is.na(chm_ge_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Gene Expression",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_cn_plot_cgc <- Heatmap(chm_cn_cgc, name = "Copy Number log2(CN/2)",
        col = colorRamp2(c(min(chm_cn_cgc[!is.na(chm_cn_cgc)]), 0, max(chm_cn_cgc[!is.na(chm_cn_cgc)])), c("white", "thistle2", "mediumorchid4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Copy Number",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_mut_plot_cgc <- Heatmap(chm_mut_cgc, name = "Mutation Status",
        na_col = "black",
        col = c("cyan3", "darkorchid"),
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Mutation Status",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)

@@

chm_ge_plot_cgc <- Heatmap(chm_ge_cgc, name = "Gene Expression (RPKM)",
        col = colorRamp2(c(min(chm_ge_cgc[!is.na(chm_ge_cgc)]), 0, max(chm_ge_cgc[!is.na(chm_ge_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Gene Expression",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_cn_plot_cgc <- Heatmap(chm_cn_cgc, name = "Copy Number",
        col = colorRamp2(c(min(chm_cn_cgc[!is.na(chm_cn_cgc)]), 0, max(chm_cn_cgc[!is.na(chm_cn_cgc)])), c("dodgerblue4", "white", "firebrick4")),
        na_col = "black",
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Copy Number",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)
chm_mut_plot_cgc <- Heatmap(chm_mut_cgc, name = "Mutation Status",
        na_col = "black",
        col = c("cyan3", "darkorchid"),
        cluster_rows = FALSE, cluster_columns = FALSE,
        row_names_gp = gpar(fontsize = 10), show_column_names = FALSE,
        column_title = "Mutation Status",
        column_title_gp = gpar(fontsize = 60, fontface = "bold"),
        column_order = chm_plot_cgc_cols,
        row_order = chm_plot_cgc_rows,
        width = 1)

@@

```{r fig.height=12, fig.width=12, include=FALSE}
makeCRISPRgrob("KRAS")
```

@@

```{r fig.height=12, fig.width=12, include=FALSE}
makeCRISPRlinplot("KRAS")
```

@@
